---
title: "Recognition memory tests analyses"
output:
  pdf_document:
    fig_caption: true
    keep_tex: yes
---

```{r global_options, include=FALSE}

# Initialization
rm(list=ls()) # Clear all

# Load all packages 
library(knitr)
library(lme4)
library(pander)
library(xtable)

# Source special purpose utilities
source('utils.R')

# General set up
use.cached.results <- TRUE # Set to FALSE to re-generate all results 
cache.directory <- '../cache'
cached.results.filename <- 'recognition.memory.logistic.glmer.results.Rda'
recognition.data.filename <- 'experiment_brisbane_recognition_memory_tests.csv'
recall.data.filename <- 'experiment_brisbane_recall_memory_tests_results.csv'

# Global chunk options
opts_chunk$set(echo=FALSE, 
               warning=FALSE, 
               message=FALSE)

# Get data 
Df.recognition <- get.recognition.data(file.path(cache.directory, recognition.data.filename))
Df.recall <- get.recall.data(file.path(cache.directory, recall.data.filename))

```

```{r general.descriptives, results='hide'}
# This is just for the calculation of subject totals, etc.
Df.session <- rbind(select(Df.recognition, session, subject, age, sex, slide),
                    select(Df.recall, session, subject, age, sex, slide)) %>% distinct() 

n.recognition.sessions <- length(unique(Df.recognition$slide))
n.recall.sessions <- length(unique(Df.recall$slide))
n.sessions <- length(unique(Df.session$slide))

n.recognition.subjects <- length(unique(Df.recognition$subject))
n.recall.subjects <- length(unique(Df.recall$subject))
n.subjects <- length(unique(Df.session$subject))

assert_that(n.recognition.sessions +  n.recall.sessions == n.sessions);

sessions.per.subject <- as.numeric(Df.session %>% 
                                     group_by(subject) %>% 
                                     summarise(count = n()) %>% 
                                     select(count) %>% 
                                     summarize(mean = mean(count)))

gender.counts.df <- Df.session %>% 
  select(subject, sex) %>% 
  distinct() %>% 
  group_by(sex) %>% 
  summarize(count = n())

gender.counts <- gender.counts.df$count
names(gender.counts) <- gender.counts.df$sex

percentile.range.lims <- function(x, p=0.95, k=1){
  q <- (1 - p)/2 
  quants <- quantile(x, probs = c(q, 1-q))
  quants[k]
}

.age.stats <- Df.session %>% 
  distinct(subject, age) %>% 
  summarize(median = median(age), 
            lower = percentile.range.lims(age, k=1), 
            upper = percentile.range.lims(age, k=2))

age.stats <- as.integer(.age.stats[1,])
names(age.stats) <- names(.age.stats)

```

In experiment *Brisbane*, each participant performed up to three separate memory tests.  In each test, the participant read a text and then had their memory tested either by a recognition or by a recall test. There were `r n.subjects` participants (`r gender.counts['Female']` female, `r gender.counts['Male']` male; the median age was `r age.stats['median']`, with 95% of participants having ages between `r age.stats['lower']` and `r age.stats['upper']`) and `r n.sessions` separate memory test sessions in total in the experiment. Of the `r n.sessions` separate tests, `r n.recognition.sessions` were recognition memory tests (performed by `r n.recognition.subjects` participants), and `r n.recall.sessions` were recall memory tests (performed by `r n.recall.subjects` participants). In general, the average number of tests done per participant was `r round(sessions.per.subject, 1)` sessions.

# Recognition memory test analysis 

## Accuracy 

```{r recognition.memory.descriptives}
hit.stats <- Df.recognition %>% 
  group_by(slide) %>%
  summarize(hits = n()) %>% 
  summarize(hit_rate = mean(hits)/20)

response.type.stats <- Df.recognition %>% 
  group_by(expected) %>%
  summarize(response = mean(response))

accuracy.stats <- Df.recognition %>%
  summarize(accuracy = mean(correct))

hit.rate <- as.numeric(hit.stats[1, 'hit_rate'])
accuracy <- as.numeric(accuracy.stats[1, 'accuracy'])

response.type.rates <- unlist(response.type.stats[,'response'])
names(response.type.rates) <- unlist(response.type.stats[,'expected'])

```

In the recognition tests, on each trial, subjects were presented with a word that was either present or absent in the text. They given 5 seconds to make their response, and across all test sessions, responses were made on `r 100*round(hit.rate,2)`% of trials. The average accuracy was `r 100*round(accuracy, 2)`%. The false positive rate was `r 100*round(response.type.rates['FALSE'], 2)`%, while the false negative rate was `r 100*round(1 - response.type.rates['TRUE'], 2)`%.

```{r }
quantile.cut <- function(x, K=50){
  breaks = quantile(x, probs=seq(0.00, 1.00, length.out=K), 
                      na.rm=T)

  cut(x, breaks=breaks)
}

transform <- function(p, eps=0.01) {
  eps/2 + (1-eps)*p 
}

logit <- function(p, eps=0.01) {
  q <- transform(p, eps) 
  log(q/(1-q))
}

Df.recognition.1 <- Df.recognition %>% mutate(pp.psi = as.numeric(scale(posterior.predictions)),
                                              cc.psi = as.numeric(scale(cooccurrence.predictions)),
                                              aa.psi = as.numeric(scale(association.predictions))) %>% 
  mutate(pp.psi.cut = quantile.cut(pp.psi),
         aa.psi.cut = quantile.cut(aa.psi),
         cc.psi.cut = quantile.cut(cc.psi))

logistic.curve.fit <- function(Df, var, title){
  
  var.cut <- paste(var, '.cut', sep='')
  
  grouping.variables <- lapply(c(var.cut, 'expected'),
                               as.symbol)
  
  Df.tmp <- Df %>%
    group_by_(.dots = grouping.variables) %>% 
    summarise_(response = 'mean(response, na.rm=T)',
               psi = sprintf('mean(%s)', var)) 
  
  eps = 0.05
  Df.tmp.2 <- Df.tmp %>%
    as.data.frame() %>% 
    select(psi, expected, response)

  M <- glm(as.formula(sprintf('response ~ %s * expected', var)), 
           family = binomial,
           data = Df)
  
  Df$predictions <- predict(M, type='response')
  Df$psi <- Df[var]
  
  ggplot(Df.tmp, 
         mapping=aes(x = psi, y = response, colour=expected)) + 
    geom_point() +
    geom_line(Df, mapping = aes(x=psi, y=predictions, colour=expected)) + 
    geom_jitter(Df, mapping = aes(x=psi, y = -.2 + .1*expected + 1.3*response, colour=expected), size=0.01, height=0.01) +
    ggtitle(title, subtitle = sprintf('AIC: %2.2f', AIC(M))) +
    scale_y_continuous(breaks = c(0.0, 0.25, 0.5, 0.75, 1.0)) + 
    theme_classic()
  
}
```


```{r}
logistic.curve.fit(Df.recognition.1, 'pp.psi', 'Topic model')
logistic.curve.fit(Df.recognition.1, 'cc.psi', 'Cooccurrence model')
logistic.curve.fit(Df.recognition.1, 'aa.psi', 'Association model')

```




```{r}

glmer.recognition.memory.analysis <- function(Df){
  
  glmer.control <- glmerControl(optimizer="bobyqa",
                                optCtrl=list(maxfun=100000))

  psi.predictors <- c('posterior.predictions',
                      'association.predictions',
                      'cooccurrence.predictions')
  
  Df[, psi.predictors] <- lapply(Df[, psi.predictors],
                                 scale)
  
  model.formula.1 <- 'response ~ expected + %s +  (1|text) + (1|subject) + (1|word)'
  model.formula.2 <- 'response ~ expected + %s +  (%s + expected|text) + (%s + expected|subject) + (1|word)'
  model.formula.3 <- 'response ~ expected + %s +  (%s + expected|text) + (%s|subject) + (1|word)'
  model.formula.4 <- 'response ~ expected * %s +  (%s + expected|text) + (%s|subject) + (1|word)'
  model.formula.5 <- 'response ~ expected * %s +  (%s + expected|text) + (%s + expected|subject) + (1|word)'
  model.formula.6 <- 'response ~ expected * %s +  (%s * expected|text) + (%s * expected|subject) + (1|word)'
    
  make.formula <- function(formula.string, var){
    as.formula(sprintf(formula.string, var, var, var))
  }
  
  logistic.glmer <- function(formula, predictor){
    withWarnings(
      glmer(make.formula(formula, predictor),
            data=Df,
            control = glmer.control,
            family=binomial)
    )
  }

  model.formulae <- list(model.1 = model.formula.1,
                         model.2 = model.formula.2,
                         model.3 = model.formula.3,
                         model.4 = model.formula.4,
                         model.5 = model.formula.5,
                         model.6 = model.formula.6)
  
  predictors <- list(TopicModel = 'posterior.predictions',
                     CooccurrenceModel = 'cooccurrence.predictions',
                     AssociationModel = 'association.predictions',
                     NullModel = '1')
                     
  lapply(model.formulae,
         function(formula){
           lapply(predictors,
                  function(predictor){
                    logistic.glmer(formula, predictor)
                  })
         }
  )
  
}
  
get.results.table <- function(model.results){
  
  results <- rbind(sapply(model.results, function(result) BIC(result$value)), 
                   sapply(model.results, function(result) AIC(result$value)), 
                   sapply(model.results, function(result) -2*logLik(result$value)[1]))
  
  rownames(results) <- c('BIC', 'AIC', 'Deviance')
  
  results
}
  

if (!use.cached.results) {
  model.results <- glmer.recognition.memory.analysis(Df.recognition)
  save('model.results',
       file=file.path(cache.directory, cached.results.filename))
} else {
  load(file.path(cache.directory, cached.results.filename))
}

```

## Model 1

```{r}
pander(get.results.table(model.results[['model.1']]))
```
## Model 2

```{r}
pander(get.results.table(model.results[['model.2']]))
```

## Model 3

```{r}
pander(get.results.table(model.results[['model.3']]))
```


## Model 4

```{r}
pander(get.results.table(model.results[['model.4']]))
```


## Model 5

```{r}
pander(get.results.table(model.results[['model.5']]))
```

## Model 6

```{r}
pander(get.results.table(model.results[['model.6']]))
```
# Bayesian data analysis

```{r}
model.names <- list(model.1 = "mcmc_recognition_model_v1_seed_666_burn_10000_sample_10000_dic_10000_thin_10.Rda",
                    model.2 = "mcmc_recognition_model_v2_seed_667_burn_10000_sample_10000_dic_10000_thin_10.Rda",
                    model.3 = "mcmc_recognition_model_v3_seed_1667_burn_10000_sample_10000_dic_10000_thin_10.Rda",
                    model.4 = "mcmc_recognition_model_v4_seed_10667_burn_10000_sample_10000_dic_10000_thin_10.Rda",
                    model.5 = "mcmc_recognition_model_v5_seed_11667_burn_10000_sample_10000_dic_10000_thin_10.Rda",
                    model.6 = "mcmc_recognition_model_v6_seed_14667_burn_10000_sample_10000_dic_10000_thin_10.Rda")

bda.results <- lapply(model.names,
                      function(model.name) readRDS(file.path(cache.directory, model.name))
)



get.max.psrf <- function(model.name){

  max(sapply(lapply(bda.results[[model.name]]$S,
                    function(s) gelman.diag(s, multivariate = F)),
             function(x) max(x$psrf[,2])))

}

get.dic <- function(model.name){

  deviance <- sapply(bda.results[[model.name]]$D,
                     function(d) round(sum(d$deviance) + sum(d$penalty))
  )

  deviance['TopicModel'] <- deviance['pp']
  deviance['CooccurrenceModel'] <- deviance['cc']
  deviance['AssociationModel'] <- deviance['aa']
  deviance['NullModel'] <- deviance['null']

  deviance[c('TopicModel', 'CooccurrenceModel', 'AssociationModel', 'NullModel')]
}

get.akaike.weights <- function(model.name){
  deviances <- get.dic(model.name)
  f <- exp(-(deviances - min(deviances))/2)
  f <- f[c('TopicModel', 'CooccurrenceModel', 'AssociationModel')]
  f/sum(f)
}

make.results.table <- function(table.func){
  do.call(rbind,
          lapply(list(model.1 = 'model.1',
                      model.2 = 'model.2',
                      model.3 = 'model.3',
                      model.4 = 'model.4',
                      model.5 = 'model.5',
                      model.6 = 'model.6'),
                 table.func)
  )
}

bda.results.table <- make.results.table(get.dic)
akaike.weights.table <- make.results.table(get.akaike.weights)

pander(bda.results.table)



```

## Akaike weights
```{r}
pander(akaike.weights.table)
```

## Recall results
```{r}
get_recall_results <- function(mcmc_results_filename){
  bda_results <- readRDS(file.path(cache.directory, 
                                   mcmc_results_filename))
}

get_recall_models_dic <- function(bda_results){
  
  get_dic <- function(bda_results, model_name){
    sum(bda_results$D[[model_name]]$deviance) + sum(bda_results$D[[model_name]]$penalty)
  }
  
  lapply(list('TopicModel' = 'pp', 
              'CooccurrenceModel' = 'cc', 
              'AssociationModel' = 'aa'),
         function(model_name) get_dic(bda_results, model_name))
  
}

recall_results <- get_recall_results('mcmc_recall_model_seed_1401_burn_1000_sample_1000_dic_1000_thin_10.Rda')
pander(unlist(get_recall_models_dic(recall_results)))
```

